---
title: "CLRS 2016 Reserve Example"
author: "SFR Actuarial Consulting, Inc."
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: false
    lightbox: false
---

```{r setup, include=FALSE}
# Sets global chunck options
knitr::opts_chunk$set(echo = TRUE)

proj_info <- list(
  clms = "claims and claim expenses",
  client = "CLRS Insurance Company",
  clnt = "CLRS IC", 
  val_date = as.Date('1997-12-31'),
  acct_date = as.Date('1997-12-31'),
  info_date = as.Date('1997-12-31')
)

auto_dat_cas <- 'http://www.casact.org/research/reserve_data/ppauto_pos.csv'

total_row <- function(df, sum_totals, total_label = NULL){
  
  # df <- summary_df
  # sum_totals <- 2:5
  # total_label <- 1
  totals <- sapply(X = df[,sum_totals], FUN = sum)
  df <- rbind(df, NA)
  
  if(!is.null(total_label)) {
    df[nrow(df), total_label] <- "Total"}
  
  df[nrow(df), sum_totals] <- lapply(df[-nrow(df), sum_totals], sum)
  
  return(df)
}


```

## Executive Summary

SFR consulting was retained by the CLRS 2016 R Workshop participants to prepare this analysis of unpaid claims and claims expenses. In this document, we present a analysis of unpaid claims and claim expenses. We developed this analysis using `R` and used R Markdown to create this document.

The relevant dates for our analysis are as follows:  

* Accounting date: `r proj_info$acct_date`  
* Valuation date: `r proj_info$val_date`  
* Information date: `r proj_info$info_date`   

The scope of our analysis is to estimate the unpaid claims and claims expenses for the automobile liability coverage for `r proj_info$client`.

### Findings

### Acknowledgement

Any questions related to this report should be directed to:

Rajesh Sahasrabuddhe, FCAS, MAAA  
(215) 246-1028  
<rajesh_saharabuddhe@oliverwyman.com>

Brian A. Fannin, ACAS 

<BFannin@RedwoodsGroup.com>  

Emma Ran Li  
(814) 753-0321  
<ril5077@gmail.com>


## Setting up the R Environment

We used several R packages to prepare this analysis. 
```{r message = FALSE, warning=FALSE, results='hide'}
# load packages

my_packages <- list("dplyr", "magrittr", "lubridate", "htmlTable", "knitr", 
  "xtable", "ChainLadder", "scales", "highcharter")
sapply(X = my_packages, FUN = library, character.only = TRUE)



```

## Data

The data underlying this analysis was provided by the Casualty Actuarial Society (CAS) which was available at <`r auto_dat_cas`>. We downloaded the data on `r format(Sys.time(), "%A, %b %d, %Y at %X")`.


```{r cache=TRUE, echo=FALSE}

auto_dat <- read.csv(file = auto_dat_cas) %>% 
  tbl_df()

# Standardize names

names(auto_dat)[names(auto_dat) == 'AccidentYear'] <- 'acc_yr'
names(auto_dat)[names(auto_dat) == 'DevelopmentYear'] <- 'dev_yr'
names(auto_dat)[names(auto_dat) == 'DevelopmentLag'] <- 'lag'
names(auto_dat)[names(auto_dat) == 'IncurLoss_B'] <- 'incur_loss'
names(auto_dat)[names(auto_dat) == 'CumPaidLoss_B'] <- 'paid_loss'
names(auto_dat)[names(auto_dat) == 'BulkLoss_B'] <- 'bulk_loss'
names(auto_dat)[names(auto_dat) == 'EarnedPremDIR_B'] <- 'dir_prem'
names(auto_dat)[names(auto_dat) == 'EarnedPremCeded_B'] <- 'ceded_prem'
names(auto_dat)[names(auto_dat) == 'EarnedPremNet_B'] <- 'net_prem'
names(auto_dat)[names(auto_dat) == 'PostedReserve97_B'] <- 'posted_reserve'
names(auto_dat) <- tolower(names(auto_dat))

model_dat <- auto_dat %>% 
  dplyr::filter(grname == 'State Farm Mut Grp')


```



We sumamrize that data in the table below:

```{r echo=FALSE}
summary_df <- model_dat %>% 
  dplyr::filter(dev_yr == year(proj_info$val_date)) %>% 
  dplyr::select(acc_yr, dir_prem, incur_loss, paid_loss) %>% 
  dplyr::mutate(case_loss = incur_loss - paid_loss)

summary_df <- total_row(summary_df, sum_totals =  2:5, total_label = 1)

summary_df[,2:5] <- lapply(summary_df[,2:5], txtInt)

htmlTable::htmlTable(summary_df, 
  caption = paste("Data Summary\nas of", proj_info$val_date),
  rnames = FALSE, 
  header = c("Year", "Premium", "Incurred", "Paid", "Case"),
  css.total = "border-top: 1px solid #BEBEBE; font-weight: 900; 
  border-bottom: 1px solid #BEBEBE",
  css.cell = "padding-left: 15px",
  align = c("c", "r", "r", "r", "r"), 
  ctable = FALSE,
  total = TRUE)


```


## Loss Development
### Paid Loss Development Experience
We present paid loss development experience in the triangle below.  

```{r echo=FALSE}

my_triangles = list(
  paid = 
    as.triangle(Triangle = dplyr::filter(model_dat, dev_yr <= 1997),
      origin = "acc_yr", dev = "lag", value = "paid_loss"),
  incurred = 
    as.triangle(Triangle = dplyr::filter(model_dat, dev_yr <= 1997),
      origin = "acc_yr", dev = "lag", value = "incur_loss")
)

y_at <- 2000000 * 1:5

plot(my_triangles$paid, lattice = TRUE, main = "Paid Loss Development",
  xlab = "Matrurity (Years)", 
  scales = list(
    y = list(at = y_at,
      labels = comma(y_at))
    ))

```

### Incurred Loss Development Experience
We present incurred loss development experience in the triangle below:
```{r echo=FALSE}

plot(my_triangles$incurred, lattice = TRUE, main = "Incurred Loss Development",
  xlab = "Matrurity (Years)",
  scales = list(
    y = list(at = y_at,
      labels = comma(y_at))
    ))

```

### Obseravtions and Comments
In our review of the development experience, we noted the follwing:  

- Paid development patterns have been generally lengthening particulalry at intermediate maturities such as 2:3 and 3:4
- Incurred development experience indicates that a neagtive IBNR is most likely appropriate. More recent accident years have exhibited more favaorable incurred development that older accident years.   

## Ultimate Loss Projections
We used the `R` package `ChainLadder` to project ultimate losses using the following methods:  

- Mack-Chain-Ladder Model applied to


### Mack-Chain-Ladder Model

```{r mcl_paid}
mcl_paid <- MackChainLadder(Triangle = my_triangles$paid)
mcl_paid
plot(mcl_paid)

```


```{r mcl_incurred}
mcl_incurred <- MackChainLadder(Triangle = my_triangles$incurred)
mcl_incurred
plot(mcl_incurred)

```

### Bootstrap-Chain-Ladder Model
```{r bcl_paid}
bcl_paid <- BootChainLadder(Triangle = my_triangles$paid, R = 999, process.distr = "gamma")
bcl_paid
plot(bcl_paid)
```

```{r bcl_incurred}
bcl_incurred <- BootChainLadder(Triangle = my_triangles$incurred, R = 999, process.distr = "gamma")
bcl_incurred
plot(bcl_incurred)
```

### Munich-Chain-Ladder Model
```{r munich}
munich <- MunichChainLadder(Paid = my_triangles$paid, Incurred = my_triangles$incurred, 
  est.sigmaP = 0.1, est.sigmaI = 0.1)
munich
plot(munich)

```
## Comparison of Methods

We now compare estimates by method


```{r}
ultimates <- data_frame(acc_yr = row.names(summary(bcl_incurred)$ByOrigin),
  bcl_incurred = summary(bcl_incurred)$ByOrigin$'Mean Ultimate',
  bcl_paid = summary(bcl_paid)$ByOrigin$'Mean Ultimate',
  mcl_incurred = summary(mcl_incurred)$ByOrigin$'Ultimate',
  mcl_paid = summary(mcl_paid)$ByOrigin$'Ultimate',
  munich_incurred = summary(munich)$ByOrigin$'Ult. Incurred',
  munich_paid = summary(munich)$ByOrigin$'Ult. Paid',
  incurred = summary(munich)$ByOrigin$'Latest Incurred',
  paid = summary(munich)$ByOrigin$'Latest Paid')


ultimates
```

```{r, echo=FALSE, message=FALSE}
hc <- highchart() %>% 
  hc_title(text = "Estimated Ultimate Claims") %>% 
  hc_xAxis(categories = ultimates$acc_yr) %>% 
  hc_plotOptions(column = list(stacking = "normal", zoomType = 'x',
            panning = TRUE, panKey = 'shift')) %>% 
  hc_plotOptions(line = list(zoomType = 'x',
            panning = TRUE, panKey = 'shift')) %>% 
  hc_add_series(data = ultimates$bcl_incurred - ultimates$paid,
      name = "Case", type = "column") %>% 
  hc_add_series(data = ultimates$paid, name = "Paid", type = "column") %>% 
  hc_add_series(data = ultimates$bcl_incurred, name = "Bootstrap Incurred", type = "line") %>% 
  hc_add_series(data = ultimates$bcl_paid, name = "Bootstrap Paid", type = "line") %>% 
  hc_add_series(data = ultimates$munich_incurred, name = "Munich Chain Ladder Incurred", type = "line") %>% 
  hc_add_series(data = ultimates$munich_paid, name = "Munich Chain Ladder Paid", type = "line") %>% 
  hc_add_series(data = ultimates$mcl_paid, name = "Mack Chain Ladder Paid", type = "line") %>% 
   hc_add_series(data = ultimates$mcl_paid, name = "Mack Chain Ladder Paid", type = "line") %>% 
  hc_tooltip(valueDecimals = 0)


hc

```





## R packages
We used the following `R` packagaes to develop this analysis:  

- `r citation()$textVersion` - 
- `r citation('ChainLadder')$textVersion`