---
title: "CLRS 2016 Reserve Example"
author: "SFR Actuarial Consulting, Inc."
date: "May 15, 2016"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r echo=FALSE}
library(dplyr)
library(magrittr)
library(lubridate)
library(htmlTable)
library(knitr)
library(xtable)
```

```{r cache=TRUE, echo=FALSE}
auto_dat_cas <- 'http://www.casact.org/research/reserve_data/ppauto_pos.csv'
auto_dat <- read.csv(file = auto_dat_cas) %>% 
  tbl_df()
```

```{r echo=FALSE}
names(auto_dat)[names(auto_dat) == 'AccidentYear'] <- 'acc_yr'
names(auto_dat)[names(auto_dat) == 'DevelopmentYear'] <- 'dev_yr'
names(auto_dat)[names(auto_dat) == 'DevelopmentLag'] <- 'lag'
names(auto_dat)[names(auto_dat) == 'IncurLoss_B'] <- 'incur_loss'
names(auto_dat)[names(auto_dat) == 'CumPaidLoss_B'] <- 'paid_loss'
names(auto_dat)[names(auto_dat) == 'BulkLoss_B'] <- 'bulk_loss'
names(auto_dat)[names(auto_dat) == 'EarnedPremDIR_B'] <- 'dir_prem'
names(auto_dat)[names(auto_dat) == 'EarnedPremCeded_B'] <- 'ceded_prem'
names(auto_dat)[names(auto_dat) == 'EarnedPremNet_B'] <- 'net_prem'
names(auto_dat)[names(auto_dat) == 'PostedReserve97_B'] <- 'posted_reserve'
names(auto_dat) <- tolower(names(auto_dat))

model_dat <- auto_dat %>% 
  dplyr::filter(grname == 'State Farm Mut Grp')

proj_info <- list(
  clms = "claims and claim expenses",
  client = "CLRS Insurance Company",
  clnt = "CLRS IC", 
  val_date = as.Date('1997-12-31'),
  acct_date = as.Date('1997-12-31'),
  info_date = as.Date('1997-12-31')
  )

```



## Executive Summary

SFR consulting was retained by the CLRS 2016 R Workshop participants to prepare this analysis of unpaid claims and claims expenses. In this document, we present a analysis of unpaid claims and claim expenses. We developed this analysis using `R` and used R Markdown to create this document.

The relevant dates for our analysis are as follows:
* Accounting date: 
* Valuation date:
* Information date

The scope of our analysis is to estimate the unpaid claims and claims expenses for the automobile liability coverage for `r proj_info$client`.

### Findings

### Acknowledgement

Any questions related to this report should be directed to:

Rajesh Sahasrabuddhe, FCAS, MAAA  
(215) 246-1028  
<rajesh_saharabuddhe@oliverwyman.com>

Brian A. Fannin, ACAS  
<BFannin@RedwoodsGroup.com>  

Emma Ran Li  
(814) 753-0321  
<ril5077@gmail.com>


## Data

The data underlying this analysis was provided by the Casualty Actuarial Society (CAS) which was available at the <`r auto_dat_cas`> on `r format(Sys.time(), "%A, %b %d, %Y at %X")`.

We sumamrize that data in the table below:

```{r echo=FALSE}
model_dat %<>% 
  dplyr::filter(dev_yr == year(proj_info$val_date)) %>% 
  dplyr::select(acc_yr, dir_prem, incur_loss, paid_loss) %>% 
  dplyr::mutate(case_loss = incur_loss - paid_loss) %>% 
  dplyr::mutate(acc_yr = as.character(acc_yr))

model_dat %<>% 
    bind_rows(
    data_frame(acc_yr = "Total", 
      dir_prem = sum(model_dat$dir_prem),
      incur_loss = sum(model_dat$incur_loss),
      paid_loss = sum(model_dat$paid_loss),
      case_loss = sum(model_dat$case_loss))) %>% 
  as.data.frame()

model_dat[,2:5] <- lapply(X = model_dat[,2:5], function(x) {
  prettyNum(x, big.mark = ",")})

htmlTable::htmlTable(model_dat, 
  caption = paste("Data Summary\nas of", proj_info$val_date),
  rnames = FALSE, 
  header = c("Year", "Premium", "Incurred", "Paid", "Case"),
  css.total = "border-top: 1px solid #BEBEBE; font-weight: 900; 
  border-bottom: 1px solid #BEBEBE",
  css.cell = "padding-left: 15px",
  align = c("c", "r", "r", "r", "r"), 
  ctable = FALSE,
  total = TRUE)






```


